---
title: "Winning Characteristics in the Olympics"
author: "Noah Obuya and Tamya Davidson"
format: pdf
editor: visual
---

```{r load-packages}
library(tidyverse)
library(tidymodels)
library(formatR)
library(MASS)
library(nnet)
library(car)
library(lme4)
library(glmnet)


```

```{r read-in-data,results = FALSE}

olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')

```

```{r}

olympics <- olympics %>% 
 na.omit

olympics04 <- olympics %>% 
  filter(year == 2004)

olympics08 <- olympics %>% 
  filter(year == 2008)

olympics12 <- olympics %>% 
  filter(year == 2012)

olympics16 <- olympics %>% 
  filter(year == 2016)
  
```

# Introduction and Data

## Research Question

**What are the most influential characteristics (between height and weight
when it comes to predicting gold medals?

The data that we chose was olympics data from TidyTuesday\'s github repository (https://github.com/rfordatascience/tidytuesday/b
lob/master/data/2021/2021-07-27/readme.md). The dataset was created in 
May 2018, and the data were collected by scraping www.sports-reference.com.  
The data contains 271,116 observations of 15 variables. The variables of 
interest in our research include sex, age, height , weight , noc (country)
, year , season, and medals (Gold, Silver, Bronze). Based on these variables 
, we will answer the question of what are the most influential variables that
influence an athlete receiving a gold medal, and if these variables change 
over time. From the dataset we will only observe the more recent olympic
games (including the years 2004, 2008, 2012, 2016). As part of our 
data-cleaning process, we have created 4 different dataset subsections 
for each of these years. Additionally, there were many NA values 
corresponding to medals. Since this is our variable of interest,
we will drop all NA values corresponding to medals. After doing 
this we are left with a case study of 39,783 observations of 15 variables. 
The motivation behind this project is to analyze what 
athletes can do to better prepare for the Olympic games, and see
which factors are more influential than others. 

\

## Variables of Interest

Variables of Interest

Sex - Sex Assigned at Birth of the Olympian

Age - Age of the Olympian in years

Height - Height of the Olympian in centimeters (cm)

Weight - Weight of the Olympian in kilograms (kg)

NOC - Country as assigned by the National Olympic Committee

##  Methodology 

We will be fitting an ordinal model. We are doing this because the outcome 
variable, Medal, is not only categorical, but ordered. Within the three 
separate categories of medal, we can assume that an Olympian who received 
a gold medal has done better than an Olympian who received a silver medal. 
The same can be said for a silver medal vs. a bronze medal and a gold medal 
vs. a bronze medal. 

The assumptions of an ordinal model include the following:

1.The dependent variable are ordered

2.One or more of the independent variables are either continuous, 
categorical or ordinal.

3.There is no multicollinearity.

4.There are proportional odds.

The first two assumptions are already taken care of through the nature of
our model. It is clear that our dependent variable, Medal, is ordered. We 
have continuous and categorical independent variables (sex and country
being categorical, height, weight and age being continuous). For the 
multicollinearity assumption, we will be examining correlation plots
between each variable and Variance Inflation Factor (VIF) tests to 
ensure multicollinearity does not exist. 

In cleaning our dataset, we removed all the NA values corresponding to 
our variable, Medal. There were NA values corresponding to our Medal 
variable because data were Missing at Random (MAR). We know that, for
each event, there can only be three placements: Gold, Silver and Bronze.
This is our observed data, because we have a variable corresponding to 
the event. Our missing data are all the NA values for medal that we 
initially had in our dataset. Therefore our missing data are related 
to observed data, because we know those who didn\'t receive a medal 
within their events (NA value) just weren\'t able to place within the
three placements that are offered for each event.

We fit an ordinal model with no transformations or interactions and
tested the assumptions to see how it would perform as an initial model; 
something we could base our improvements off of. This ordinal model 
included our variables of interest, with Medals as our outcome. We first 
ran LASSO on this model, selecting important variables from the output
LASSO gave. Then, we ran all-subset selection on the variables that were
chosen by our LASSO output. We opted not to use forward selection or 
backward elimination because we knew our variables were highly correlated,
which we cover in the discussion of our transformations and interactions.

\
The predictor variables we are considering for the model are sex, age, 
height, weight and NOC (country). We considered interactions terms between
sex \* weight and sex \* height. 

\

## Exploratory Data Analysis

```{r}
olympics04 %>%
  count(medal) %>%
  mutate(per = n/sum(n)) %>%
  ggplot()
```

```{r}
olympics08 %>%
  count(medal)%>%
  mutate(per = n/sum(n))
```

```{r}
olympics12 %>%
  count(medal) %>%
  mutate(per = n/sum(n))
```

```{r}
olympics16 %>%
  count(medal) %>%
  mutate(per = n/sum(n))
  
```

In 2004, the number of bronze medals handed out to individuals was 676 which 
was 33.8% of the total medals, the number of silver medals was 660 which was 
33% of the total medals , and the number of gold medals was 664 which was 33.2%
of the total medals .

In 2008, the number of bronze medals handed out to individuals was 706 which
was 34.7% of the total medals , the number of silver medals was 665 which was
32.7% of the total medals, and the number of gold medals was 664 which was 
32.6% of the total medals.

In 2012, the number of bronze medals handed out to individuals was 669 which
was 35% of the total medals , the number of silver medals was 624 which was 
32.6% of the total medals, and the number of gold medals was 622 which was
32.4% of the total medals.

In 2016, the number of bronze medals handed out to individuals was 700 which
was 34.8% of the total medals, the number of silver medals was 652 which was
32.3% of the total medals, and the number of gold medals was 662 which was 
32.8% of the total medals.

```{r}
ggplot(olympics , mapping = aes(x = weight , y = sex )) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(x = "Weight", y = "Sex", title = "Distribution of weights of 
       athletes by sex", subtitle = "Men have a higher median weight 
       than women across all olympic games")
```

As we can see from the boxplots above, the distribution of weight for men and 
\women athletes competing in the olympics are both skewed to the right, while 
it appears that the men are skewed heavier. We were interested in the one 
female athlete who is considered an outlier because her weight is above 150.
We have found the athlete to be Olha Vasylivna Korobka who actually got a 
silver medal in the 2008 summer games in weight lifting. (Code shown below).

```{r}
olympics %>%
  filter(sex == "F") %>%
  filter(weight > 150)
```

Height vs Sex BoxPlots

```{r}
ggplot(olympics , mapping = aes(x =height , y = sex )) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(x = "Height", y = "Sex", title = "Distribution of heights of athletes 
       by sex", subtitle = "Men have a higher median height than women across
       all olympic games")
```

Similar to the results that we saw in the boxplots comparing distributions of 
weights between men and women, we can see that men also have a higher median
height than women who have completed in the Olympics.

*We chose to use all years when analyzing the distribution of heights and
weights* because over the course of our time frame (2004 - 2016) there have 
been many rule changes about allowed and not allowed substances, 
and analyzing these two variables through all of the years can give us 
a better idea of distributions.

Now that we have analyzed the data and got some idea of the distribution 
of specific parameters of interest, we are interested in analyzing which 
variables are the biggest factor in predicting gold medals for Summer 
Olympic games, and whether or not these variables (and their influence) 
change over time.

## Logistic Regression

First we will fit a logistic regression model that predicts the probability
of receiving a gold medal (for the purpose of this model, we will use the 
goldMedal? column that gives us a 1 if someone received a gold medal for 
their event, and gives us a 0 if someone did not receive a gold medal 
(they received either gold or silver) this is due to the characteristics
of logistic regression and how it works best when predicting a binary outcome.)

```{r,echo=FALSE}
olympics_goldMedal <- olympics %>%
  mutate(goldMedal = if_else(medal == "Gold", 1, 0)) 
```

```{r,echo=FALSE}
 logit_mod <- glm(goldMedal ~ sex  + age + height +  weight + noc, 
                  data = olympics_goldMedal, family = "binomial"(link = "logit")) 
   tidy(logit_mod) %>%
     mutate(estimate = round(estimate, 9))
```

The expected log odds of someone achieving a gold medal if their sex is male
is 0.0535 times higher than if someone is a female when holding all other
variables constant. For every one year increase in age, the expected log 
odds of someone achieving a gold medal is expected to increase by .00164 
when all other variables are held constant. For every one unit increase
in height, we expect the logs odds of someone achieving a gold medal to 
increase by approximately 0.0020 when all other variables are held constant.
. For every one unit increase in weight, we expect the log odds of someone
achieving a gold medal to increase by approximately 0.00022 when all other
variables are held constant. For each respective noc, the expected log
odds of someone achieving a gold medal to \[increase or decrease\] by X 
when all other variables are held constant.

## Ordinal Regression

```{r,echo=FALSE}
olympics_ord <- 
  olympics%>%
  mutate(medals = case_when(
    medal == "Gold"~ 3, 
                  medal == "Silver"~2, 
           medal == "Bronze"~1, 
           TRUE ~ 0
  ))
ordMod <-
  polr(factor(medals) ~ sex  + age + height +  weight + noc, 
       data = olympics_ord , method = "probit")
tidy(ordMod) %>%
  mutate(estimate = round(estimate , 9))
```

```{r,results = FALSE}
exp(coef(ordMod))
```

## 

## Variable Selection

```{r,results = FALSE}
y <- olympics_ord$medals
x <- model.matrix(medals ~ sex  + age + height +  weight + noc ,
                  data = olympics_ord)
m_lasso_cv <- cv.glmnet(x, y, alpha = 1)
```

```{r}
best_lambda <- m_lasso_cv$lambda.min
best_lambda
```

```{r}
plot(m_lasso_cv)
```

```{r, results = FALSE}
m_best <- glmnet(x, y, alpha = 1, lambda = best_lambda)
m_best$beta
```

```{r,results = FALSE}
library(leaps)
m_all_noc <- regsubsets(medals ~ sex  + age + height +  weight + noc,
                  data = olympics_ord, 
                  nbest = 1, nvmax = 5, really.big = T)
m_all_noc

```

```{r,results = FALSE}
summary(m_all_noc)
```

```{r}
m_all <- regsubsets(medals ~ sex  + age + height +  weight,
                  data = olympics_ord, 
                  nbest = 1, nvmax = 5, really.big = T)
m_all

```

```{r}
summary(m_all)
```

```{r}
summary(m_all)$cp
```

Final Model

```{r}
ordMod_final <-
  polr(factor(medals) ~ sex  + age + height , data = olympics_ord , method = "probit")
tidy(ordMod_final) %>%
  mutate(estimate = round(estimate , 9))

```

## Results

LASSO removed weight as a variable and majority of all NOC identifiers (country
) as variables as well. When we ran our all subset selection on these variables,
we got back the same data, which removed weight and noc, leaving sex, age and 
height.

##  Discussion

In summary, we learned that weight does not have an impact on whether an
Olympian receives a gold medal or not. In addition, we learned that when 
taking into consideration our variables of interest, sex age and height 
are the only influential variables when we perform variable selection.\
For our future work, we could subset this data by years and see if the 
influence of these variables change over time. Additionally, we could check 
the variable\'s influence with Cook\'s Distance.

\
