---
title: "Winning Characteristics in the Olympics"
author: "Noah Obuya and Tamya Davidson"
format: pdf
editor: visual
---

## Read Data and Packages

```{r laod-packages, warning = FALSE, message = FALSE, tidy = TRUE, tidy.opts=list(width.cutoff=25)}

library(tidyverse)
library(tidymodels)
install.packages("formatR")
library(formatR)
library(MASS)
library(nnet)
install.packages("car")
library(car)
library(lme4)
```

```{r read-in-data, warning = FALSE, message = FALSE, tidy = TRUE, tidy.opts=list(width.cutoff=25)}

olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')

```

## Brainstorm 

What variable in our model is most influential (or most accurately) able to predict someone winning a gold medal?

What country has won the most gold medals?

If someone doesn't win a gold medal, are they more likely to win a silver or bronze medal?

Do men or women win more gold medals , and are the influential variables the same for them?

*Are countries with warmer climates more prone to winning a gold medal during the summer olympics?* Would have to find some data classifying a country as a warm country.
