---
title: "Winning Characteristics in the Olympics"
author: "Noah Obuya and Tamya Davidson"
format: pdf
editor: visual
---

```{r load-packages, warning = FALSE, message = FALSE, tidy = TRUE, echo = FALSE, tidy.opts=list(width.cutoff=25)}

library(tidyverse)
library(tidymodels)
install.packages("formatR")
library(formatR)
library(MASS)
library(nnet)
install.packages("car")
library(car)
library(lme4)
```

```{r read-in-data, warning = FALSE, message = FALSE, tidy = TRUE, echo = FALSE, tidy.opts=list(width.cutoff=25)}

olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')

```

## Brainstorm

What variable in our model is most influential (or most accurately) able to predict someone winning a gold medal?

What country has won the most gold medals?

If someone doesn't win a gold medal, are they more likely to win a silver or bronze medal?

Do men or women win more gold medals , and are the influential variables the same for them?

*Are countries with warmer climates more prone to winning a gold medal during the summer olympics?* Would have to find some data classifying a country as a warm country.

## Filtering Data

```{r new data sets, warning = FALSE, message = FALSE, tidy = TRUE, echo = FALSE, tidy.opts=list(width.cutoff=25)}

olympics <- olympics %>% 
 na.omit

olympics04 <- olympics %>% 
  filter(year == 2004)

olympics08 <- olympics %>% 
  filter(year == 2008)

olympics12 <- olympics %>% 
  filter(year == 2012)

olympics16 <- olympics %>% 
  filter(year == 2016)
  
```

```{r playing around, warning = FALSE, message = FALSE, tidy = TRUE, echo = FALSE, tidy.opts=list(width.cutoff=25)}
test <- olympics %>%
  mutate(old = if_else(age > 25, "Old", "Young")) %>%  
  filter(year == 2004 | year == 2008 | year == 2012 | year == 2016)%>%  
  filter(!is.na(medal)) %>%
  filter(medal == "Gold") %>%
  count(noc) %>%
  filter(n > 50)
```

# Introduction and Data

## Research Question

## Variables of Interest

## Data Cleaning

## Exploratory Data Analysis

# Methodology

## Why We Chose an Ordinal Model

## Model Selection Process

## Variable Interactions and Transformations

## Model Conditions and Diagnostics

# Results

## Final Model

## Key Results and Interpretations

# Discussion

## Revisiting the Research Question

## Hypothesis Test

## Limitations

## Ideas for Future Work

# Appendix
